Type-safe client-server communication for web applications

### 2014-02-28 paper.tex
Add related work for defunctionalization and cloud haskell.

@@

@@

@@

>>>>

\paragraph{Inspiration and alternatives to \lstinline!export!} One crucial
aspect of implementing cross-network function calls is, of course, the issue of
data representation: the client side of things must be able to obtain some
representation of any function it may want to call. In our solution, this
representation is obtained through the use of the \lstinline!export! function,
which when executed on the server pairs a function with a unique identifier and
on the client, and when executed on the client returns said identifier so that
the client may now refer to the function. While this has the advantage of being
simple to implement, one major drawback of this method is that all functions
must be explicitly exported in the \lstinline!App! monad prior to being called
over the network.

This approach was inspired by Cloud Haskell \cite{cloudhaskell}, which
introduces the notion of ``static values''; values which are known at compile
time. Codifying this in the type system, to enable it to be used as a basis for
remote procedure calls, unfortunately requires some major changes to the
compiler. Cloud Haskell has a stopgap measure for unmodified compilers wherein
a remote table, pairing values with unique identifiers, are kept. This explicit
bookkeeping relies on the programmer to assign appropriate types to both values
themselves and their identifiers, breaking type safety.

The astute reader may notice that this is exactly what the \lstinline!export!
does as well, the difference being that \export!export! links the identifier to
the value it represents on the type level, making it impossible to call
non-existent remote functions and break the program's type safety in other ways.

Another approach to this problem is defunctionalization
\cite{defunctionalization}, a program transformation wherein functions are
translated into algebraic data types. This approach would allow the client and
server to use the same actual code; rather than passing an identifier around,
the client would instead pass the actual defunctionalized function. This would
have the added benefit of allowing functions to be arbitrarily composed before
being remotely invoked.

This approach also requires significant changes to the compiler, making it
unsuitable for our use case. Moreover, we are not entirely convinced about the
wisdom of allowing server-side execution of what is essentially arbitrary code
sent from the client which, for web applications, is completely untrustworthy.

@@

\bibitem[Danvy, Nielsen (2001)]{defunctionalization}
O. Danvy, L. R. Nielsen: Defunctionalization at work, 2001

@@

\bibitem[Epstein et al (2011)]{cloudhaskell}
J. Epstein, A. P. Black, S. Peyton Jones: Towards Haskell in the Cloud, 2011

@@

\bibitem[Svenningsson, Svensson (2013)]{bjornbenny}
J. Svenningsson, J. Svensson: Simple and Compositional Reification of Monadic Embedded Languages, 2013
